---
title: "Rimerman-Final"
author: "Mitch Rimerman"
date: "2022-12-07"
output: pdf_document
---

```{r setup, include=FALSE, results=FALSE, message=FALSE, warning=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(reticulate)
library(tidyverse)
library(viridis)
use_virtualenv("py-proj")
```


```{python}
import numpy as np
import torch 
import torchvision
import matplotlib.pyplot as plt
```

## Part I

Given a matrix $u\in\mathbb{R}^{m\times n}$, the following code provides functions for row-wise and column-wise backward difference of $u$, $v=dxb(u)$ and $v=dyb(u)$ respectively.

```{python}
def dxb(u):
  return u-np.roll(u,1,axis=0)

def dyb(u): 
  return u-np.roll(u,1,axis=1)

```

The following code provides functions for row-wise and column-wise forward difference of $u$, $v=dxf(u)$ and $v=dyf(u)$ respectively.

```{python}
def dxf(u):
  return np.roll(u,-1,axis=0)-u

def dyf(u): 
  return np.roll(u,-1,axis=1)-u

```


\newpage
## Part II

The original image:
```{python, out.width="70%"}
griffith=torchvision.io.read_image('griffith.jpg')
griffith=torchvision.transforms.Resize((400,400))(griffith)
plt.imshow(griffith.permute(1, 2, 0))
plt.show()
```

The grayscale image:

```{python, out.width="70%"}
griffith_gray = torchvision.transforms.Grayscale(num_output_channels=1)(griffith)
griffith_gray = torch.reshape(griffith_gray, (400,400))
plt.imshow(griffith_gray, cmap='Greys_r')
plt.show()
```

\newpage
The grayscale image with Gaussian noise:

```{python, out.width="70%"}
noise = torch.zeros(400,400, dtype=torch.float64)
noise = noise + (16)*torch.randn(400, 400)
griffith_gn = griffith_gray+noise

griffith_gray_np = griffith_gray.numpy()
griffith_gn_np = griffith_gn.numpy()

plt.imshow(griffith_gn, cmap='Greys_r')
plt.show()
```


\newpage
## Part III
### A

```{python}
def matrix_dif_normed(u):
  eps = 2**(-52)
  uxf = dxf(u)
  uyf = dyf(u)
  denom = np.sqrt(uxf*uxf+uyf*uyf+eps**2)
  p1 = uxf/denom
  p2 = uyf/denom
  return np.array([p1,p2])

def Df(u,f,mu):
  p = matrix_dif_normed(u)
  D = dxb(p[0])+dyb(p[1])
  return -D+mu*(u-f)

def GD(f,mu,alpha,iterations):
  u=np.zeros((400,400))
  for i in range(iterations):
    u_next = u - alpha*Df(u,f,mu)
    u = u_next
  plt.imshow(u, cmap='Greys_r')
  plt.show()
```

\newpage
### B

For the following code, due to the inverse relationship between $\alpha$ and iteration, I chose $\alpha=.01$ and iterations$=1000$. To first find a the range of values for $\mu$ which produce the best images, I test $\mu=\{10^{-3},10^{-2},10^{-1}, 1/2, 1, 10\}$ 


```{python, fig.show="hold", out.width="33%"}
alpha = .01
iterations = 1000
mus  = [.001,.01,.1,.5,1,10]

f=griffith_gn_np
for mu in mus:
  u=np.zeros((400,400))
  for i in range(iterations):
    u_next = u - alpha*Df(u,f,mu)
    u = u_next
  plt.imshow(u, cmap='Greys_r')
  plt.show()

```

From these images, we see that the best value for $\mu$ is $0.1$. Thus, the optimal range for $\mu$ is on the order of $10^{-1}$.

\newpage
Now, to further optimize $\mu$, I test a range of values close to $0.1$, namely $\mu=\{.025,.05,.075,.1,.125,.15\}$ and examine the images produced. Here, I also compare the reproduced image to the original image numerically using image error, to help with the choice of $\mu$.


```{python, fig.show="hold", out.width="33%"}
alpha = .01
iterations = 1000

mus  = [.025,.05,.075,.1,.125,.15]
error = []

f=griffith_gn_np
for mu in mus:
  u=np.zeros((400,400))
  for i in range(iterations):
    u_next = u - alpha*Df(u,f,mu)
    u = u_next
  error.append((1/160000)*sum(sum(((u-griffith_gray_np)**2))))
  plt.imshow(u, cmap='Greys_r')
  plt.show()

print(f'{error[0:3]}\n{error[3:7]}')
```

Visually, the best images are produced with $\mu=0.05$ and $\mu=0.075$ Although the image error decreases as we increase $\mu$, there appears to be much more noise in the image.

```{python}
print((1/160000)*sum(sum(((f-griffith_gray_np)**2))))
```

If we look at the error between the original image and the noisy image, we see that the error here is very low, but it is clearly a noisy image, so image error is not the best metric to use when selecting $\mu$. Thus, we will proceed with $\mu=0.065$

\newpage
### C

Now, we will increase the number of iterations to get an even better image. We continue with $\mu=0.065$ and $\alpha=.01$, and complete $10,000$ iterations.

```{python, out.width="70%"}
alpha = .01
iterations = 10000
mu = .065
choice_error = []

f=griffith_gn_np
u=np.zeros((400,400))
for i in range(iterations):
  u_next = u - alpha*Df(u,f,mu)
  u = u_next
  choice_error.append((1/160000)*sum(sum(((u-griffith_gray_np)**2))))
  
plt.imshow(u, cmap='Greys_r')
plt.show()

```

\newpage
### D

```{r, out.width="70%"}
c_error <- unlist(py$choice_error)
c_error_df <- data.frame(1:10000,c_error)
names(c_error_df) <- c("Iteration","Error")
ggplot(c_error_df) +
  geom_line(aes(x=Iteration,y=Error), color="#2D4790", size=.75)+
  labs(title="Image Error Per Iteration")+
  ylim(0,25000)+
  theme_minimal()
```


